Neural Connectome of the Ctenophore Statocyst

### 2025-05-01 manuscript/Manuscript.qmd
MS M&M

Larvae of *Mnemiopsis leidyi* (five days old) were cryofixed using a high-pressure freezing apparatus (HPM Live μ, CryoCapCell, https://www.cryocapcell.com/technologies/high-pressure-freezer) and immediately transferred to liquid nitrogen for storage. The frozen samples were processed in a substitution medium containing 2% (w/v) osmium tetroxide and 0.5% uranyl acetate in acetone, using a cryo-substitution device (EM AFS-2, Leica). Cryo-substitution was carried out by gradually raising the temperature  <!-- steps? solutions? --> , and the samples were embedded in epoxy resin <!--type? -->. Serial untrithin sections of 50 nm thickness were prepared using a Reichert Jung Ultracut E ultramicrotome and a 45º DiATOME diamond knife <!-- wasn't it our Leica UC7? -->. To enhance section adhesion and improve hydrophilicity, a conductive indium tin oxide-coated glass slide (ITO Glass, UQG Optics) was treated with air glow discharge using the PELCO easiGlow system (Ted Pella, Inc.), rendering the carbon film <!-- was there carbon film? --> surface negatively charged. Section ribbons were collected on the prepared glass slides, slowly dried to ensure proper stretching, and firmly adhered to the glass surface. The sections on glass were stained with UranyLess and lead citrate (Reynolds) using airless staining bottles (Delta Microscopies). The glass slides were mounted on STEM-specific stages (Zeiss) using Copper Foil EMI Shielding Tape (3M). Imaging was done on a Gemini SEM 500 (Zeiss) equipped with SmartSEM and Atlas 5 imaging software (Zeiss). The full dataset consisted of 620 serial sections, each 50 nm thick. The imaging resolution was 2.8 nm/pixel, using electron holography transmission <!-- is this true? --> at an acceleration voltage of 1.5 kV (in-lens detector, dwell time: 3 µs  <!-- sounds very low -->).

@@

For ciliary imaging, we used the cydippid stage of *M. leidyi* at five days post-fertilization. We gently placed a coverslip with a thin layer of Vaseline applied to its two edges on a slide glass. Filtered natural seawater was introduced under the coverslip, along with the cydippids. By gently moving the coverslip, we could adjust the orientation of the larva, and once positioned, we applied slight pressure to immobilize the cydippids. We imaged the movement of balancer cilia with a differential interference contrast (DIC) microscope (Zeiss Axio Imager.M2) equipped with a highly sensitive CMOS monochrome camera optimized for the near-infrared (NIR) range (UI-3360CP-NIR-GL Rev.2, iDS) and a 40x glycerine-immersion objective lens (Objective LD LCI Plan-Apochromat 40x/1.2 Imm Corr DIC M27). To stabilize the statolith's position, we tilted the microscope 90 degrees, arranging the stage vertically. For recording, we employed a custom-made NIR LED strobe illumination system (wavelength: 850 nm) synchronized with the camera's exposure signals. The camera and LED strobe system were operated using the Video Capture Software BohNavi. Images were recorded at a resolution of 640×480 pixels with a frame rate of 100 fps for 2 minutes, using 0.05 ms pulses from the LED. To analyze the ciliary beating of the balancer cilia, we utilized the Multi Kymograph plugin in Fiji.

>>>>

Larvae of *Mnemiopsis leidyi* (five days old) were cryofixed using a high-pressure freezing apparatus (HPM Live μ, CryoCapCell, https://www.cryocapcell.com/technologies/high-pressure-freezer) and immediately transferred to liquid nitrogen for storage. The frozen samples were processed in a substitution medium containing 2% (w/v) osmium tetroxide and 0.5% uranyl acetate in acetone, using a cryo-substitution device (EM AFS-2, Leica). Cryo-substitution was carried out by gradually raising the temperature from -90°C to -70°C over 4 hours, then returning to -20°C over 2 hours, and finally to room temperature over 2 hours. The samples were then embedded in epoxy resin (EMbed 812, Electron Microscopy Sciences). Serial ultrathin sections of 50 nm thickness were prepared using a Leica UC7 ultramicrotome and a 45º DiATOME diamond knife. To enhance section adhesion and improve hydrophilicity, a conductive indium tin oxide-coated glass slide (ITO Glass, UQG Optics) was treated with air glow discharge using the PELCO easiGlow system (Ted Pella, Inc.), rendering the surface negatively charged. Section ribbons were collected on prepared glass slides, gently dried to ensure proper stretching, and firmly adhered to the glass surface. The sections were stained with UranyLess and lead citrate (Reynolds) using airless staining procedures. The glass slides were mounted on a STEM-specific stage (Zeiss) using Copper Foil EMI Shielding Tape (3M). Imaging was performed using a Gemini SEM 500 (Zeiss), equipped with SmartSEM and Atlas 5 imaging software (Zeiss), using the Inlens detector. A total of 620 serial sections, each 50 nm thick, were analyzed. The imaging resolution was 2.8 nm/pixel, and the acceleration voltage was set to 1.5 kV. The imaging time for each tile was 54 minutes and 21 seconds, with a pixel size of 2.8 nm, a tile size of 91.8 × 91.8 µm (32768 × 32768 pixels), and a dwell time of 3.0 µs.

@@

For ciliary imaging, we used the cydippid stage of *M. leidyi* at five days post-fertilization. A coverslip with a thin layer of Vaseline applied to its two edges was gently placed over a glass slide. Filtered natural seawater and the cydippids were introduced beneath the coverslip. The orientation of the larvae was adjusted by carefully sliding the coverslip, and slight pressure was then applied to immobilize the larvae.

Balancer cilia movement was imaged using a differential interference contrast (DIC) microscope (Zeiss Axio Imager.M2) equipped with a high-sensitivity monochrome CMOS camera optimized for the near-infrared range (UI-3360CP-NIR-GL Rev.2, iDS), and a 40x glycerine-immersion objective lens (LD LCI Plan-Apochromat 40x/1.2 Imm Corr DIC M27). To stabilize the statolith, the microscope was tilted by 90° to achieve vertical stage alignment. Illumination was provided by a custom-made NIR LED strobe system (850 nm) synchronized with the camera’s exposure signals. Image acquisition was controlled using BohNavi software. Recordings were acquired at 640 × 480 pixels, 100 frames per second (fps), for a duration of 2 minutes, using 0.05 ms pulse illumination from the LED.

For beat frequency analysis, kymographs were generated using the Multi Kymograph plugin in Fiji. One-second segments were manually extracted from the kymographs, and ciliary beat frequencies were counted visually for each ROI. These manually determined frequencies were then used for subsequent computational analysis.

To process the beat frequency time series data, left and right cilia traces were extracted, filtered to remove missing or non-finite values, and assigned frame indices. Data were saved as RDS objects for downstream analysis. Mean values were plotted as bar graphs with the left and right cilia shown in positive and negative directions, respectively.

To assess coordination between left and right cilia, we computed rolling Pearson correlation coefficients over a window of 20 frames. For each dataset, we excluded windows with insufficient valid data or no variation. Time series of correlation coefficients were plotted and summarized as mean values. These mean correlations were then compared across planes (sagittal vs. tentacular) using Welch's t-test. The resulting statistics were visualized as boxplots with jittered data points.
